---
title: "Music Data Project"
author: "Thomas Schechter"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
#Start by loading packages

library(pacman)

p_load(readr,data.table,tidyverse,ggplot2,tidymodels,ranger,parsnip,purrr,magrittr,lubridate,parallel,pbapply,stringr)

#We'll add more as needed
```
```{r}
#Make sure you commit to your repo

#After that, it's time to load in datasets

lyrics_data <- "~/data_work/music_data/lyrics-data.csv" %>% read.csv()
artists_data <-"~/data_work/music_data/artists-data.csv" %>% read.csv()
music_album_reviews <- "~/data_work/music_data/music_album_reviews.csv" %>% read.csv()
music_genre <- "~/data_work/music_data/music_genre.csv" %>% read.csv() 

#Imported directly from the directory on the bottom-right
```

```{r}
#Peruse the data, get a feel for it all. It's a lot...

glimpse(lyrics_data)

lyrics_sum <- summary(lyrics_data)

lyrics_sum

#It's all character data in lyrics_data

artists_sum <- artists_data %>% summary()

artists_sum

#3/5 is character, 2/5 is numerical
```

```{r}
#continue onto genre and album review datasets

genre_sum <- music_genre %>% summary()

reviews_sum <- music_album_reviews %>% summary()

genre_sum

#some character, mostly numerical data
```

```{r}

reviews_sum

#1 column characters (the review), 1 column numeric (The rating)

#Interesting that 4.0 is the first quartile in review rating... people like what they listen to?

#mean is somewhat below the median for this data set. Let's check it out.
```
```{r}
#let's do some visualization
 #First, simplify the name of the dataset

rev_data <- music_album_reviews

#find the length of each review. We're going to plot that against the rating each album recieved in the review

rev_data$rev_length <- str_length(rev_data$Review)

rev_data %>% glimpse()

rating_chart <- ggplot(rev_data, aes(x=rev_length,y=Rating, color=rev_length)) + geom_boxplot() + facet_wrap(vars(Rating))


rating_chart
```
```{r}
#Let's do some cleaning. 
#Impute to remove any NAs with the mean of the Ratings

rev_data$Rating[is.na(rev_data$Rating)] <- mean(rev_data$Rating, na.rm=T)

anyNA(rev_data$Rating)

#We now can see we have no missing values and can use the data with assurance it's clean

rev_data$bad_review <- ifelse(rev_data$rev_length<=140,print("TRUE"),print("FALSE"))
                              
view(rev_data)
```

```{r}

```


```{r}
#How many cores are we working with?

detectCores(all.tests = FALSE, logical = TRUE)

#12. Cool. Let's use 10

cls <- makeCluster(10)


```






MUSIC GENRES

```{r}
#Let's make this easier to understand

genre_data <- music_genre

summary(genre_data)

#Let's change the tempo to numeric data

genre_data$tempo <- genre_data$tempo %>% as.numeric()

#Don't forget to eliminate missing values (impute)

genre_data$tempo[is.na(genre_data$tempo)] <- mean(genre_data$tempo, na.rm=T)

anyNA(genre_data$tempo)

#SUCCESS
```
```{r}
#What is the most represented genre in the dataset?

genre_data$electr <- ifelse(genre_data$music_genre=="Electronic","TRUE","FALSE")

view(genre_data)

#See what other genres are here, and repeat for all

#unique(genre_data$music_genre)

#genre_data$blues <- ifelse(genre_data$music_genre=="Blues","TRUE","FALSE")
#genre_data$anime <- ifelse(genre_data$music_genre=="Anime","TRUE","FALSE")
#genre_data$rock <- ifelse(genre_data$music_genre=="Rock","TRUE","FALSE")
#genre_data$alt <- ifelse(genre_data$music_genre=="Alternative","TRUE","FALSE")
#genre_data$country <- ifelse(genre_data$music_genre=="Country","TRUE","FALSE")
#genre_data$classical <- ifelse(genre_data$music_genre=="Classical","TRUE","FALSE")
#genre_data$rap <- ifelse(genre_data$music_genre=="Rap","TRUE","FALSE")
#genre_data$hiphop <- ifelse(genre_data$music_genre=="Hip-Hop","TRUE","FALSE")

#view(genre_data)

#Now it's easier to find the number of rows and we have more options


#Above is entirely unnecessary... just use this below
elect_num <- nrow(filter(genre_data,music_genre=="Electronic"))
blues_num <- nrow(filter(genre_data,music_genre=="Blues"))
classic_num <- nrow(filter(genre_data,music_genre=="Classical"))
rock_num <-nrow(filter(genre_data,music_genre=="Rock"))
anime_num <- nrow(filter(genre_data,music_genre=="Anime"))
rap_num <- nrow(filter(genre_data,music_genre=="Rap"))
hiphop_num <- nrow(filter(genre_data,music_genre=="Hip-Hop"))
country_num <-nrow(filter(genre_data,music_genre=="Country"))
alt_num <- nrow(filter(genre_data,music_genre=="Alternative"))

elect_num
blues_num
hiphop_num
rock_num
rap_num
country_num
alt_num
anime_num
classic_num


#It seems there's 5k of each genre. It probably said this somewhere in the data description, but I missed that
```

```{r}
#visualize?

genre_plot1 <- ggplot(genre_data, aes(x=energy,y=popularity, color=music_genre)) + geom_point() + facet_wrap(vars(key)) 


genre_plot1

#Wow... this is cluttered.

```
```{r}
#Create a few new variables to filter it down.

summary(genre_data)

#clean and impute anything that needs it

genre_data$popularity[is.na(genre_data$popularity)] <- mean(genre_data$popularity, na.rm=T)
genre_data$acousticness[is.na(genre_data$acousticness)] <- mean(genre_data$acousticness, na.rm=T)
genre_data$danceability[is.na(genre_data$danceability)] <- mean(genre_data$danceability, na.rm=T)
genre_data$liveness[is.na(genre_data$liveness)] <- mean(genre_data$liveness, na.rm=T)
genre_data$energy[is.na(genre_data$energy)] <- mean(genre_data$energy, na.rm=T)
genre_data$instrumentalness[is.na(genre_data$instrumentalness)] <- mean(genre_data$instrumentalness, na.rm=T)
genre_data$loudness[is.na(genre_data$loudness)] <- mean(genre_data$loudness, na.rm=T)
genre_data$speechiness[is.na(genre_data$speechiness)] <- mean(genre_data$speechiness, na.rm=T)
```

```{r}
#Explore a bit with regression analysis and plotting

#Does a faster tempo indicate a louder song?

lm_tempo_loud <- lm(tempo ~ loudness, data=genre_data)
lm_loud_tempo <- lm(loudness ~ tempo, data=genre_data)

lm_tempo_loud
lm_loud_tempo
```
It appears that tempo and loudness (dynamics) are positively correlated.

$Tempo = 1.07*Loudness+129.73$

$Loudness = 0.04809*Tempo -14.90246$

Perhaps these musicians don't know how to play fast softly, or their audience just doesn't want that? Let's find that out...

```{r}
#Regression of lm_tempo_loud on popularity

lm_tl_pop <- lm(popularity ~ tempo + loudness, data=genre_data)

lm_tl_pop

#try with an interaction term?

lm_tl_pop2 <- lm(popularity ~ tempo + loudness + tempo*loudness, data=genre_data)


lm_tl_pop2

```
It seems louder music leads to more listening. Faster music? Not so much. Louder and Faster? Not a chance.


```{r}
#Make some new variables...

genre_data$dummy_pop <- ifelse(genre_data$popularity>50,1,0)

view(genre_data)

#Convert to a data table to make it easier to play with

genre_data <- as.data.table(genre_data)

#Clean up the table a bit

genre_data <- genre_data[,c(1:18,80)]
view(genre_data)

#Find out how many song fit our "popular" criteria (Popular > 50)

genre_data %>% filter(dummy_pop==1) %>% nrow()

18900*3

#A little over a third, it seems, are "popular"
```

```{r}
#Let's do a breakdown of the popular

#Create a new data set for the popular songs

popular_songs <- genre_data[dummy_pop==1]

summary(popular_songs)
```

```{r}
#Clean it up a bit

popular_songs <- popular_songs[,!c("dummy_pop")]

summary(popular_songs)

#Make some more datasets to play with, don't get too messy
```

```{r}
pop_rock <- popular_songs[music_genre=="Rock"]
pop_classic <- popular_songs[music_genre=="Classical"]
pop_anime <- popular_songs[music_genre=="Anime"]
pop_rap <- popular_songs[music_genre=="Rap"]
pop_country <- popular_songs[music_genre=="Country"]
pop_hiphop <- popular_songs[music_genre=="Hip-Hip"]
pop_elec <- popular_songs[music_genre=="Electronic"]
pop_alt <- popular_songs[music_genre=="Alternative"]
pop_blues <- popular_songs[music_genre=="Blues"]

pop_rock %>% summary()

pop_rock_fast <- pop_rock %>% filter(tempo > 140)

pop_rock_fast_plot <- ggplot(pop_rock_fast, aes(x=popularity, y=energy, color=key)) + geom_point()

pop_rock_slow <- pop_rock %>% filter(tempo<101.19)

pop_rock_slow_plot <- ggplot(pop_rock_slow, aes(x=popularity, y=energy, color=key)) + geom_point()

pop_rock_fast_plot

pop_rock_slow_plot
```

The most popular rock songs don't seem to depend on tempo very much.

```{r}
pop_rock_slow %>% nrow()

pop_rock_fast %>% nrow()
```

There are almost twice as many fast popular rock songs than there are slow... interesting, but we did limit the slower ones below the first quartile while the faster ones were the whole quartile... Perhaps we can run a regression to see more.


```{r}
lm_pop_rock_tempo <- lm(popularity ~ tempo, data=pop_rock)

lm_pop_rock_tempo
```
EXPLORING KEYS

Interestingly, it seems tempo has a negative effect on popularity. Why is this? Let's explore with keys, whether minor or major.

Minor keys are more sorrowful and dark, while majors tend to be brighter and happier.

```{r}
#First we will explore major vs. minor before diving more in depth.

#Create a dummy, since major vs minor is binary

pop_rock$dummy_major <- ifelse(pop_rock$mode=="Major",1,0)

view(pop_rock)
```

```{r}
#With a dummy created, regress

lm_major <- lm(popularity ~ dummy_major, data=pop_rock)

tidy(lm_major)
```

It doesn't seem that picking between a major or minor key has a terribly significant effect on popularity...

Shall we try with some specific keys?

```{r}
#dummy for keys to try

pop_rock$dummy_amin <- ifelse(pop_rock$key=="A" & pop_rock$mode=="Minor",1,0)

#create a linear model

lm_amin <- lm(popularity ~ dummy_amin, data=pop_rock)

lm_amin %>% tidy()
```

Not even significant in the slightest!!!

Let's try another key.

```{r}
# Dummy for Eb Major

pop_rock$dummy_emaj <- ifelse(pop_rock$key=="E" & pop_rock$mode=="Major",1,0)

lm_emaj<- lm(popularity~dummy_emaj, data=pop_rock)
  
lm_emaj %>% tidy()
```

Still insignificant statistically. Let's try something else: minor key and in C


```{r}
#Dummy for C major OR minor


pop_rock$dummy_c <- ifelse(pop_rock$key=="C",1,0)

lm_c <- lm(popularity~dummy_c,data=pop_rock)

lm_c %>% tidy()

#Insignificant. What about C major?

lm_c_maj <- lm(popularity~dummy_c+dummy_major+dummy_c*dummy_major, data=pop_rock)

lm_c_maj %>% tidy()
```
Nothing significant here.


```{r}
#Let's visualize a few of these variables now...

#Tempo and popularity?

plot_tempo_pop <- ggplot(pop_rock, aes(x=tempo,y=popularity,color=key)) + geom_point()


```


